On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions

### 2013-09-27 paper2.tex

%MNRAS
\title[On the Physics of Radio Halos in Galaxy Clusters]{On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2 *}$, Christoph Pfrommer$^{3 \dagger}$ and Francisco Prada$^{4,5,1 \S}$\\
$^{1}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{2}$now at GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  The underlying physics of giant and mini radio halos in galaxy clusters is
  still an open question, which becomes more pressing with the growing number of
  detections. Assuming that the radio-emitting electrons are generated in
  hadronic cosmic ray (CR) interactions with protons of the intra-cluster medium
  explains the radio emission in mini halos (Perseus, Ophiuchus). {\bf By contrast,
  the hadronic model either fails to explain the extended emission of giant
  radio halos (Coma at low frequencies) or would require a flat CR profile, 
  realized through CR streaming and diffusion (Coma and A2163 at 1.4~GHz). 
  We suggest that a second, leptonic component could be responsible 
  for the missing flux in giant halos within a new hybrid scenario of which 
  we describe possible observational consequences. We then study in
  detail the hadronic component, including an effective parameterization of CR
  transport, using a cosmological complete mock galaxy cluster catalog built 
  from the MultiDark $N$-body $\Lambda$CDM simulation in a companion paper.} 
  Because of CR transport and the different scalings of the
  X-ray luminosity, $L_{\rmn{X}}$, and the Sunyaev-Zel'dovich flux, $Y$, with
  gas density, our model can simultaneously reproduce the observed bimodality of
  radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
  unimodal distribution of radio-halo luminosity versus $Y$; thereby suggesting
  a physical solution to this apparent contradiction. {\bf Assuming a redshift-independent
  fraction of radio-loud clusters, we demonstrate the unique 
  prospects for low-frequency radio surveys (such as the
  LOFAR Tier~1 survey) to detect about $600\times\rmn{\emph{loud~fraction}[\%]}/10$ 
  hadronically-generated giant and mini radio halos. 
  This shows the potentially of future radio surveys in determining 
  the generation mechanism of halos and the importance of the hadronic component in 
  galaxy clusters, particularly at low masses which are currently unexplored.}
  We make our cosmologically complete multi-frequency mock 
  catalogs for the (non-)thermal cluster emission at different redshifts publicly and freely 
  available on-line through the MultiDark database.
%  The underlying physics of giant and mini radio halos in galaxy clusters is
%  still an open question, which becomes more pressing with the growing number of
%  detections. Assuming that the radio-emitting electrons are generated in
%  hadronic cosmic ray (CR) interactions with protons of the intra-cluster medium
%  explains the radio emission in mini halos (Perseus, Ophiuchus). By contrast,
%  the hadronic model either fails to explain the extended emission of giant
%  radio halos (Coma at low frequencies) or would require a flat CR profile 
%  (Coma and A2163 at 1.4~GHz). Flat CR distributions are only expected to be realized 
%  through CR streaming in {\em relaxed} clusters---in conflict with the observation that
%  radio halos are hosted by merging {\em turbulent} clusters. This motivates us
%  to propose a new hybrid model, where (i) mini halos are primarily of hadronic
%  origin, (ii) giant halos experience a transition from central hadronic
%  emission to peripheral leptonic emission due to Fermi I or II re-acceleration
%  of fossil or hadronically produced electrons, and (iii) steep spectrum radio
%  sources are mainly of leptonic origin. {\bf We use a mock galaxy cluster catalog
%  built from the MultiDark $N$-body $\Lambda$CDM simulation that matches 
%  Sunyaev-Zel'dovich and X-ray scaling relations and luminosity function.} 
%  Because of CR streaming and the different scalings of the
%  X-ray luminosity, $L_{\rmn{X}}$, and the Sunyaev-Zel'dovich flux, $Y$, with
%  gas density, our model can simultaneously reproduce the observed bimodality of
%  radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
%  unimodal distribution of radio-halo luminosity versus $Y$; thereby suggesting
%  a physical solution to this apparent contradiction. For a plausible fraction
%  of 10\% radio-loud clusters, our model matches the NVSS radio-halo luminosity
%  function. Constructing an analytical radio-halo luminosity function, we
%  demonstrate the unique prospects for low-frequency radio surveys (such as the
%  LOFAR Tier 1 survey) to detect $\sim3500$ radio halos back to redshift two and
%  to probe the underlying physics of radio halos. We make our cosmologically 
%  complete multi-frequency mock catalogs for the (non-)thermal cluster emission 
%  at different redshifts publicly and freely available on-line
%  through the MultiDark database.
\end{abstract}

@@

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\let\thefootnote\relax\footnotetext{$\dagger$ christoph.pfrommer@h-its.org}
\let\thefootnote\relax\footnotetext{$\S$ fprada@iaa.es}
\endgroup

@@

>>>>

%MNRAS
\title[On the Physics of Radio Halos in Galaxy Clusters]{On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2}$, Christoph Pfrommer$^{3}$ and Francisco Prada$^{4,5,1}$\\
$^{1}$GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands, f.zandanel@uva.nl\\
$^{2}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany, christoph.pfrommer@h-its.org\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf The underlying physics of giant and mini radio halos in galaxy clusters
    is still an open question. We find that mini halos (such as in Perseus and
    Ophiuchus) can be explained by radio-emitting electrons that are generated
    in hadronic cosmic ray (CR) interactions with protons of the intra-cluster
    medium. By contrast, the hadronic model either fails to explain the extended
    emission of giant radio halos (as in Coma at low frequencies) or would
    require a flat CR profile, which can be realized through outward streaming
    and diffusion of CRs (in Coma and A2163 at 1.4~GHz).  We suggest that a
    second, leptonic component could be responsible for the missing flux in
    giant halos within a new hybrid scenario and we describe its possible
    observational consequences.  To study the hadronic emission component of the
    radio halo population statistically, we use a cosmological mock galaxy
    cluster catalog built from the MultiDark simulation. Because of the
    properties of CR streaming and the different scalings of the X-ray
    luminosity ($L_{\rmn{X}}$) and the Sunyaev-Zel'dovich flux ($Y$) with gas
    density, our model can simultaneously reproduce the observed bimodality of
    radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
    unimodal distribution of radio-halo luminosity versus $Y$; thereby
    suggesting a physical solution to this apparent contradiction. We predict
    radio halo emission down to the mass scale of galaxy groups, which
    highlights the unique prospects for low-frequency radio surveys (such as the
    LOFAR Tier 1 survey) to increase the number of detected radio halos by at
    least an order of magnitude.}
\end{abstract}

@@

@@

% original abstract:

%  The underlying physics of giant and mini radio halos in galaxy clusters is
%  still an open question, which becomes more pressing with the growing number of
%  detections. Assuming that the radio-emitting electrons are generated in
%  hadronic cosmic ray (CR) interactions with protons of the intra-cluster medium
%  explains the radio emission in mini halos (Perseus, Ophiuchus). By contrast,
%  the hadronic model either fails to explain the extended emission of giant
%  radio halos (Coma at low frequencies) or would require a flat CR profile 
%  (Coma and A2163 at 1.4~GHz). Flat CR distributions are only expected to be realized 
%  through CR streaming in {\em relaxed} clusters---in conflict with the observation that
%  radio halos are hosted by merging {\em turbulent} clusters. This motivates us
%  to propose a new hybrid model, where (i) mini halos are primarily of hadronic
%  origin, (ii) giant halos experience a transition from central hadronic
%  emission to peripheral leptonic emission due to Fermi I or II re-acceleration
%  of fossil or hadronically produced electrons, and (iii) steep spectrum radio
%  sources are mainly of leptonic origin. {\bf We use a mock galaxy cluster catalog
%  built from the MultiDark $N$-body $\Lambda$CDM simulation that matches 
%  Sunyaev-Zel'dovich and X-ray scaling relations and luminosity function.} 
%  Because of CR streaming and the different scalings of the
%  X-ray luminosity, $L_{\rmn{X}}$, and the Sunyaev-Zel'dovich flux, $Y$, with
%  gas density, our model can simultaneously reproduce the observed bimodality of
%  radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
%  unimodal distribution of radio-halo luminosity versus $Y$; thereby suggesting
%  a physical solution to this apparent contradiction. For a plausible fraction
%  of 10\% radio-loud clusters, our model matches the NVSS radio-halo luminosity
%  function. Constructing an analytical radio-halo luminosity function, we
%  demonstrate the unique prospects for low-frequency radio surveys (such as the
%  LOFAR Tier 1 survey) to detect $\sim3500$ radio halos back to redshift two and
%  to probe the underlying physics of radio halos. We make our cosmologically 
%  complete multi-frequency mock catalogs for the (non-)thermal cluster emission 
%  at different redshifts publicly and freely available on-line
%  through the MultiDark database.
